SwiftKey Capstone Proj. Milestone Report

### 2019-11-13 draft1.Rmd
papa loves mambo

Probs of words completing OBSERVED trigrams
```{r}
gamma2 <- 0.5  # bigram discount
gamma3 <- 0.5  # trigram discount
bigPre <- 'is_book'

## Returns a two column data.frame of observed trigrams that start with the
## bigram prefix (bigPre) in the first column named ngram and
## frequencies/counts in the second column named freq. If no observed trigrams
## that start with bigPre exist, an empty data.frame is returned.
##
## bigPre -  single-element char array of the form w2_w1 which are the first 
##           two words of the trigram we are predicting the tail word of
## trigrams - 2 column data.frame or data.table. The first column: ngram,
##            contains all the trigrams in the corpus. The second column:
##            freq, contains the frequency/count of each trigram.
getObsTrigs <- function(bigPre, trigrams) {
    trigs.winA <- data.frame(ngrams=vector(mode = 'character', length = 0),
                             freq=vector(mode = 'integer', length = 0))
    regex <- sprintf("%s%s%s", "^", bigPre, "_")
    trigram_indices <- grep(regex, trigrams$ngram)
    if(length(trigram_indices) > 0) {
        trigs.winA <- trigrams[trigram_indices, ]
    }

@@

## Returns a two column data.frame of observed trigrams that start with bigram
## prefix bigPre in the first column named ngram and the probabilities
## q_bo(w_i | w_i-2, w_i-1) in the second column named prob calculated from
## eqn 12. If no observed trigrams starting with bigPre exist, NULL is returned.
##
## obsTrigs - 2 column data.frame or data.table. The first column: ngram,
##            contains all the observed trigrams that start with the bigram
##            prefix bigPre which we are attempting to the predict the next
##            word of in a give phrase. The second column: freq, contains the
##            frequency/count of each trigram.
## bigrs - 2 column data.frame or data.table. The first column: ngram,
##         contains all the bigrams in the corpus. The second column:
##         freq, contains the frequency/count of each bigram.
## bigPre -  single-element char array of the form w2_w1 which are first two
##           words of the trigram we are predicting the tail word of
## triDisc - amount to discount observed trigrams
getObsTriProbs <- function(obsTrigs, bigrs, bigPre, triDisc=0.5) {
    if(nrow(obsTrigs) < 1) return(NULL)
    obsCount <- filter(bigrs, ngram==bigPre)$freq[1]
    obsTrigProbs <- mutate(obsTrigs, freq=((freq - triDisc) / obsCount))
    colnames(obsTrigProbs) <- c("ngram", "prob")

@@

Find UNOBSERVED trigram tail words
```{r}
## Returns a character vector which are the tail words of unobserved trigrams
## that start with the first two words of obsTrigs (aka the bigram prefix).
## These are the words w in the set B(w_i-2, w_i-1) as defined in the section
## describing the details of equation 17.
##
## obsTrigs - character vector of observed trigrams delimited by _ of the form:
##            w3_w2_w1 where w3_w2 is the bigram prefix
## unigs - 2 column data.frame of all the unigrams in the corpus:
##         ngram = unigram
##         freq = frequency/count of each unigram
getUnobsTrigTails <- function(obsTrigs, unigs) {
    obs_trig_tails <- str_split_fixed(obsTrigs, "_", 3)[, 3]
    unobs_trig_tails <- unigs[!(unigs$ngram %in% obs_trig_tails), ]$ngram
    return(unobs_trig_tails)
}

>>>>

Probs of words completing OBSERVED trigrams
```{r}
gamma2 <- 0.5  # bigram discount
gamma3 <- 0.5  # trigram discount
bigPre <- 'and_tell'

## Returns a two column data.frame of observed trigrams that start with the
## bigram prefix (bigPre) in the first column named ngram and
## frequencies/counts in the second column named freq. If no observed trigrams
## that start with bigPre exist, an empty data.frame is returned.
##
## bigPre -  single-element char array of the form w2_w1 which are the first 
##           two words of the trigram we are predicting the tail word of
## trigrams - 2 column data.table. The first column: ngram,
##            contains all the trigrams in the corpus. The second column:
##            freq, contains the frequency/count of each trigram.
getODTrigs = function(bigPre, trigrams) {
    tri.obs = data.frame(ngrams = vector(mode = 'character', length = 0),
                          freq = vector(mode = 'integer', length = 0))
    regex = sprintf("%s%s%s", "^", bigPre, "_")
    indices = grep(regex, trigrams$ngram)
    if(length(indices > 0)) {
        tri.obs = trigrams[indices, ]
    }
    return(tri.obs)
}

#mikes's code
getObsTrigs <- function(bigPre, trigrams) {
    trigs.winA <- data.frame(ngrams=vector(mode = 'character', length = 0),
                             freq=vector(mode = 'integer', length = 0))
    regex <- sprintf("%s%s%s", "^", bigPre, "_")
    trigram_indices <- grep(regex, trigrams$ngram)
    if(length(trigram_indices) > 0) {
        trigs.winA <- trigrams[trigram_indices, ]
    }

@@

## Returns a two column data.frame of observed trigrams that start with bigram
## prefix bigPre in the first column named ngram and the probabilities
## q_bo(w_i | w_i-2, w_i-1) in the second column named prob calculated from
## eqn 12. If no observed trigrams starting with bigPre exist, NULL is returned.
##
## obsTrigs - 2 column data.frame or data.table. The first column: ngram,
##            contains all the observed trigrams that start with the bigram
##            prefix bigPre which we are attempting to the predict the next
##            word of in a give phrase. The second column: freq, contains the
##            frequency/count of each trigram.
## bigrs - 2 column data.frame or data.table. The first column: ngram,
##         contains all the bigrams in the corpus. The second column:
##         freq, contains the frequency/count of each bigram.
## bigPre -  single-element char array of the form w2_w1 which are first two
##           words of the trigram we are predicting the tail word of
## triDisc - amount to discount observed trigrams
getODTProbs = function(obsTrigs, bigrs, bigPre, triDisc = 0.5) {
    #return NULL if no trigrams starting with bigPre are observed
    if(nrow(obsTrigs) <1) return(NULL)
    #get total count of the observed trigrams
    obsCount = filter(bigrs, ngram == bigPre)$freq
    obsTrigProbs = mutate(obsTrigs, freq = ((freq - triDisc)/obsCount))
    colnames(obsTrigProbs) = c("ngram", "prob")

#Mike's code
getObsTriProbs <- function(obsTrigs, bigrs, bigPre, triDisc=0.5) {
    if(nrow(obsTrigs) < 1) return(NULL)
    obsCount <- filter(bigrs, ngram==bigPre)$freq[1]
    obsTrigProbs <- mutate(obsTrigs, freq=((freq - triDisc) / obsCount))
    colnames(obsTrigProbs) <- c("ngram", "prob")

@@

Find UNOBSERVED trigram tail words
```{r}
## Returns a character vector which are the tail words of unobserved trigrams
## that start with the first two words of obsTrigs (aka the bigram prefix).
## These are the words w in the set B(w_i-2, w_i-1) as defined in the section
## describing the details of equation 17.
##
## obsTrigs - character vector of observed trigrams delimited by _ of the form:
##            w3_w2_w1 where w3_w2 is the bigram prefix
## unigs - 2 column data.frame of all the unigrams in the corpus:
##         ngram = unigram
##         freq = frequency/count of each unigram
getUODTTails = function(obsTrigs, unigs) {

}

#mike's code
getUnobsTrigTails <- function(obsTrigs, unigs) {
    obs_trig_tails <- str_split_fixed(obsTrigs, "_", 3)[, 3]
    unobs_trig_tails <- unigs[!(unigs$ngram %in% obs_trig_tails), ]$ngram
    return(unobs_trig_tails)
}
